\section{2. Source data and inherited
limitations}\label{source-data-and-inherited-limitations}

\subsection{2.1 Published spectra}\label{published-spectra}

The analysis uses the solid infrared spectroscopic ellipsometry curves
in Fig. 6 of Moazzami \emph{et al.} \citep{Moazzami2005}.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1364}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1364}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1818}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1818}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1818}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1818}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Specimen
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Published trace
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Nominal \(x\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(T\) (K)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Thickness (um)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Reconstructed points
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
MZ226 & Fig. 6a, solid IRSE & 0.226 & 300 & 15.40 & 125 \\
MZ310 & Fig. 6b, solid IRSE & 0.310 & 300 & 4.95 & 115 \\
\end{longtable}

The published \(\alpha(E)\) curves are not native direct measurements of
absorption coefficient. They are outputs of the source study's
spectroscopic-ellipsometry and optical-model analysis
\citep{Moazzami2005, GarciaCaurel2013}. The present work adds a second
inference layer by reconstructing those plotted curves and applying
alternative downstream edge-extraction models. Consequently, this study
is conditional on the source optical inversion, layer model, thickness
assumptions, and unreported covariance.

Both specimens come from one publication. Specimen-level composition
uncertainty, carrier type, carrier density, and source inversion
covariance are unavailable. The two records therefore do not constitute
independent experimental validation.

\subsection{2.2 Bitmap reconstruction}\label{bitmap-reconstruction}

The curves were recovered from embedded one-bit bitmap objects in the
primary PDF. Each panel has a linear photon-energy axis and logarithmic
absorption axis. The horizontal calibration is 0.000350877 eV per pixel.
The declared horizontal extraction half-width is 1.5 pixels,
corresponding to 0.526 meV. The maximum vertical line-width half-range
is 0.02548 in \(\log_{10}\alpha\).

The committed reconstruction uses the per-column median source-pixel
center followed by nondecreasing isotonic regression in
\(\log_{10}\alpha\); figure-derived numerical recovery is treated as a
separate reconstruction layer \citep{Burda2017}. Source-pixel minima,
maxima, center coordinates, calibration anchors, masks, sampling
interval, and file hashes are retained. The result reproduces the
visible plotted trace at the resolution of the embedded figure; it is
not an instrument export.

A pair of adjacent MZ226 source-pixel centers reverses the expected
local direction near 0.198 eV. Isotonic regression pools that violation
into a short plateau. This is a reconstruction diagnostic, not an
identified physical feature. It is not highlighted in the main figure
and is tested only through a secondary two-point omission analysis.

\subsection{2.3 Semantics of the reported
ranges}\label{semantics-of-the-reported-ranges}

The following quantities are kept separate:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Coherent calibration-corner sensitivity:} global perturbations
  of the reconstructed energy and log-absorption coordinates within the
  declared extraction half-ranges.
\item
  \textbf{Reconstruction-method sensitivity:} changes produced by
  alternative deterministic mappings from source-pixel coordinates to a
  fitted spectrum.
\item
  \textbf{Fit-domain and weighting sensitivity:} changes caused by
  declared endpoint and point-weighting conventions.
\item
  \textbf{Fitted-model dependence:} differences among fitted intercepts
  generated by different functional forms.
\item
  \textbf{Operational-coordinate dependence:} differences among
  fixed-\(\alpha\) crossings, which are not included in the fitted-model
  span.
\end{enumerate}

None of these is presented as a confidence interval, posterior
distribution, or complete experimental uncertainty budget.
