\section{4. Numerical methods}\label{numerical-methods}

\subsection{4.1 Nominal fitting}\label{nominal-fitting}

The nominal fitted population satisfies

\[
600\leq\alpha\leq5000\ \mathrm{cm}^{-1}.
\]

Each intercept is selected by deterministic search over 4,001 uniformly
spaced edge candidates. Amplitudes and, where applicable, the fractional
exponent are solved analytically at each candidate. Residuals are
evaluated in logarithmic absorption coordinates. A candidate terminating
at a declared edge-search bound is retained for completeness but labeled
boundary-limited.

The \textbf{primary headline span} excludes boundary-limited candidates.
The full span is reported only as a completeness record.

\subsection{4.2 Coherent calibration-corner
sensitivity}\label{coherent-calibration-corner-sensitivity}

Four coherent perturbation cases are evaluated:

\[
(E_i',\log_{10}\alpha_i')=
(E_i\pm\delta E,\log_{10}\alpha_i\pm\delta_i).
\]

The nominal fit membership is held fixed to isolate coordinate
displacement from fit-window membership. This test samples global
calibration corners; it is not a complete model of local digitization
error.

\subsection{4.3 Two-point reconstruction
diagnostic}\label{two-point-reconstruction-diagnostic}

The two adjacent MZ226 coordinates associated with the visible plateau
are omitted without interpolation or replacement, and every fitted model
is recomputed. The omission is also combined with the four
calibration-corner cases. This check answers only whether that local
reconstruction irregularity controls the fitted result.

\subsection{4.4 Contiguous-window
leverage}\label{contiguous-window-leverage}

Every possible contiguous one-, three-, and five-point window is omitted
from the nominal fitted population. This is an adversarial leverage
diagnostic. It is not a missing-data probability model and does not
imply that any omitted coordinate is erroneous.

\subsection{4.5 Reconstruction-method
sensitivity}\label{reconstruction-method-sensitivity}

The fitted-model ensemble is recomputed for:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  the committed isotonic reconstruction;
\item
  the raw source-pixel centerline;
\item
  independently recomputed unweighted isotonic regression;
\item
  inverse-vertical-width-weighted isotonic regression;
\item
  isotonic reconstructions of the lower and upper source-line envelopes.
\end{enumerate}

Two membership conventions are reported separately: fixed membership
inherited from the committed nominal spectrum and
reconstruction-specific membership. This prevents preprocessing effects
from being silently conflated with changes in which points enter the
fit.

\subsection{4.6 Fit-domain and weighting
grid}\label{fit-domain-and-weighting-grid}

The lower fit boundary is varied over

\[
\alpha_{\min}\in\{400,600,800,1000\}\ \mathrm{cm}^{-1},
\]

and the upper boundary over

\[
\alpha_{\max}\in\{3000,4000,5000\}\ \mathrm{cm}^{-1}.
\]

Each admissible domain is evaluated with:

\begin{itemize}
\tightlist
\item
  equal weight per reconstructed point;
\item
  quadrature weight by photon-energy interval;
\item
  quadrature weight by log-absorption interval.
\end{itemize}

The nominal \(600\)-\(5000\) cm\(^{-1}\) interval follows the
above-bandgap modeling interval declared in the source record. The
surrounding endpoint grid brackets nearby analyst choices without
entering the deep low-absorption tail or extending substantially beyond
the source-study fitting regime. Every retained domain must contain at
least 20 reconstructed coordinates and must preserve the predeclared
edge-search bounds. The latter two weighting rules test whether
nonuniform sampling density materially determines the fitted intercept.
Excluded endpoint combinations are reported rather than repaired by
changing the search bounds.

\subsection{4.7 Model-adequacy
diagnostics}\label{model-adequacy-diagnostics}

For each nominal fitted model, the following descriptive diagnostics are
reported:

\begin{itemize}
\tightlist
\item
  RMS residual in \(\log_{10}\alpha\);
\item
  maximum absolute residual in \(\log_{10}\alpha\);
\item
  RMS residual normalized by the reconstructed source-line half-width;
\item
  fraction of fitted coordinates within one and two vertical
  half-widths.
\end{itemize}

These are compatibility diagnostics, not chi-square tests. AIC, BIC,
likelihood ratios, and formal confidence levels are not used because the
bitmap-derived coordinate errors are neither independent nor assigned a
stochastic model.

For a secondary descriptive screen, a fitted form is labeled
\textbf{line-envelope compatible} only when it is not boundary-limited,
its RMS residual normalized by the local reconstructed line half-width
is no greater than 1, and at least 95\% of fitted coordinates lie within
two local half-widths. This deterministic rule is used only to show
whether the complete model envelope is being enlarged by visibly poor
fits. It is not a physical model-selection criterion, and conclusions
are reported both with and without the screen.

The screen itself is subjected to a predeclared threshold grid:
normalized-RMS limits of 0.75, 1.0, 1.25, and 1.5 are crossed with
minimum two-half-width coverage fractions of 0.90, 0.95, and 0.975. This
analysis tests whether the retained candidate set or its span is an
artifact of the nominal screen. Because the stricter combinations can
retain fewer than two candidates, the screen is not used as the primary
model-dependence result.

\subsection{4.8 Empirical-equation scale and composition
resolution}\label{empirical-equation-scale-and-composition-resolution}

Hansen and Seiler predictions are evaluated only to provide a numerical
comparison scale. No equation is labeled the winner.

For the Hansen relation,

\[
\frac{\partial E_g}{\partial x}
=1.93-1.62x+2.496x^2-2(5.35\times10^{-4})T.
\]

At 300 K, the derivative is approximately 1.370 eV per composition
fraction at \(x=0.226\) and 1.347 eV per composition fraction at
\(x=0.310\). Half of the last reported composition digit,
\(\Delta x=0.0005\), corresponds to approximately 0.685 and 0.673 meV,
respectively. Both values exceed the 0.255 and 0.177 meV Hansen-Seiler
separations. Since the actual specimen composition uncertainty is
unreported, a sub-meV material-law ranking is not identified.
