\section{Introduction}
\label{sec:introduction}

The composition- and temperature-dependent band gap of $\mathrm{Hg}_{1-x}\mathrm{Cd}_{x}\mathrm{Te}$ controls infrared cutoff wavelength, carrier statistics, transport, and the interpretation of optical and magneto-optical measurements. The Hansen--Schmit--Casselman relation remains a widely used empirical description \cite{Hansen1982}, while later work introduced nonlinear temperature dependence and alternative static terms \cite{Seiler1990,Laurenti1990}. Over technologically relevant composition ranges, predictions from these relations can differ by only a few millielectronvolts.

A comparison at that scale is meaningful only when the measured quantity is defined with comparable precision. An absorption spectrum does not contain a unique model-independent point labeled as the band gap. The reported edge may be obtained from a power-law extrapolation, a fixed absorption crossing, a Kane-region fit, a tail-corrected model, or another operational rule. Carrier filling can shift the apparent absorption edge in degenerate HgCdTe \cite{Sarusi1989}; mercury-vacancy states can alter edge behavior and compromise a conventional optical gap criterion \cite{Yue2006}. Composition error, calibration, spectral resolution, and the selected fit domain introduce additional ambiguity. A narrow optimizer uncertainty therefore does not by itself identify the latent material gap.

This work treats edge extraction as an observation-model problem rather than fitting another universal HgCdTe polynomial. We ask three questions:
\begin{enumerate}
  \item How much does the extracted edge move under admissible observation definitions on real spectra?
  \item Can that movement change the apparent ranking of established material-gap relations?
  \item What paired measurements and specimen metadata are required to separate the latent gap from observation-specific contributions?
\end{enumerate}

We answer the first two questions using calibrated infrared spectroscopic ellipsometry spectra reported by Moazzami \emph{et al.} \cite{Moazzami2005}. Every candidate, exclusion, boundary-limited result, calibration assumption, and missing metadata field is exported through a fail-closed analysis contract. The third question is addressed through a local identifiability model and an audit-grade paired acquisition design. No candidate is promoted as a corrected or production edge, and no universal replacement for Hansen is claimed.
